Information
This is an example code of a HILICPos + HILICNeg targeted, QE run.
This code retrieves mol/L from peak areas of targeted compounds.


```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE) 
library(tidyverse)
library(parallel)
options(scipen=999)
```

Get information from standards that can be quantified.

Import standards, and filter NAs.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
Ingalls.Standards <- read.csv("data_extras/Ingalls_Lab_Standards.csv") %>%
  filter(Column == "HILIC") %>%
  select(Compound.Name, z, QE.RF.ratio, HILICMix, Conc..uM, Emperical.Formula) %>%
  filter(!is.na(Conc..uM))

## Duplicate rows testing for hilicpos and hilicneg IDs.
test <- which(duplicated(Ingalls.Standards$Compound.Name))
duplicate_compounds <- as.data.frame(Ingalls.Standards$Compound.Name[test]) %>%
  rename(Compound.Name = 1)
```

Import transect files, munge, and join with standards file.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
HILICS.transect <- read.csv("data_processed/QC_Output.csv", header = TRUE) %>%
  slice(-1:-6) %>%
  select(-c(Description, Value)) %>%
  mutate(ReplicateName = as.character(ReplicateName)) %>%
  select(ReplicateName, Metabolite.name, Column, Area.with.QC, AreaValue, Run.Type)
HILICS.transect$ReplicateName <- gsub("^.{0,1}", "", HILICS.transect$ReplicateName)

HILICS.raw.transect <- HILICS.transect %>%
  filter(Metabolite.name %in% Ingalls.Standards$Compound.Name) %>%
  rename(Compound.Name = Metabolite.name)

# Apply appropriate filters and isolate standards.
HILICS.transect <- HILICS.transect %>%
  filter(Metabolite.name %in% Ingalls.Standards$Compound.Name) %>%
  filter(str_detect(ReplicateName, "Std")) %>%
  mutate(AreaValue = as.numeric(as.character(AreaValue))) %>%
  mutate(Area.with.QC = as.numeric(as.character(Area.with.QC))) %>%
  mutate(Compound.Name = Metabolite.name) %>%
  select(-Metabolite.name) %>%
  left_join(Ingalls.Standards, by = "Compound.Name") %>%
  select(ReplicateName, Compound.Name, everything())
```

********************************************
*Manual removal of duplicate compounds*
```{r echo=FALSE, error=, message=FALSE, warning=FALSE}

HILICS.transect <- HILICS.transect %>%
  group_by(Compound.Name) %>%
  mutate(pos.and.neg = ifelse(Compound.Name %in% duplicate_compounds$Compound.Name, TRUE, FALSE)) %>%
  filter(!(pos.and.neg == TRUE & z == -1)) %>%
  filter(!(Compound.Name == "Inosine" & Column == "HILICNeg")) %>%
  filter(!(Compound.Name == "Guanine" & Column == "HILICNeg")) %>%
  mutate(z = ifelse(Column == "HILICNeg", -1, 1)) %>%
  select(-pos.and.neg)
```
********************************************

Get response factors and response factor RATIOS for transect HILIC compounds.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
RFs.transect <- HILICS.transect %>%
  mutate(RF = as.numeric(as.character(Area.with.QC))/as.numeric(Conc..uM))

RFs2.transect <- RFs.transect %>%
  mutate(Type = paste(Env = ifelse(str_detect(ReplicateName, "StdsMix|InH2O"), "Standards", "Water"),
                      Matrix = ifelse(str_detect(ReplicateName, "InMatrix"), "Matrix", "Water"), sep = "_")) %>%
  filter(str_detect(ReplicateName, as.character(HILICMix)) | str_detect(ReplicateName, "H2OInMatrix")) %>%
  filter(!str_detect(Compound.Name, ",")) %>%
  mutate(ReplicateName = substr(ReplicateName, 1, nchar(ReplicateName)-2))

# Calculate RF max and min using only standards in water.
transect.RFs.dimensions <- RFs2.transect %>%
  filter(Type == "Standards_Water") %>%
  group_by(Compound.Name) %>%
  mutate(RF.max = max(RF, na.rm = TRUE),
         RF.min = min(RF, na.rm = TRUE))

transect.RFs.dimensions$RF.max[is.infinite(transect.RFs.dimensions$RF.max) | is.nan(transect.RFs.dimensions$RF.max) ] <- NA
transect.RFs.dimensions$RF.min[is.infinite(transect.RFs.dimensions$RF.min) | is.nan(transect.RFs.dimensions$RF.min) ] <- NA

transect.RFs.dimensions <- transect.RFs.dimensions %>%
  mutate(RF.diff = RF.max/RF.min) %>%
  unique()

# Calculate the response factor ratios using (Standards in Matrix - Water in Matrix) / (Standards in Water) for each replicate.
transect.RFratios <- RFs2.transect %>%
  group_by(Compound.Name, Type) %>%
  mutate(RF.mean.per_sampleID = mean(RF, na.rm = TRUE)) %>%
  select(ReplicateName, Compound.Name, Type, RF.mean.per_sampleID) %>%
  unique() %>%
  group_by(Compound.Name) %>% filter(n() >= 3) %>%
  mutate(RF.ratio = 
           ((RF.mean.per_sampleID[Type == "Standards_Matrix"] - RF.mean.per_sampleID[Type == "Water_Matrix"]) / 
              RF.mean.per_sampleID[Type == "Standards_Water"])) %>%
  select(Compound.Name, RF.ratio) %>%
  unique()

transect.RFratios$RF.ratio[transect.RFratios$Compound.Name == "Choline"] <- 1
transect.RFratios$RF.ratio[transect.RFratios$Compound.Name == "Trimethyl-L-lysine"] <- NA ## DROPPED DUE TO SUPER WEIRD NUMBERS
transect.RFratios$RF.ratio[is.nan(transect.RFratios$RF.ratio)] <- NA

rm(list = c("RFs.transect", "RFs2.transect"))

```


Replace NA RF ratios with QE RF ratios from Ingalls Standards.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
test.RFratios <- transect.RFratios %>%
  filter(is.na(RF.ratio))

test.standards <- Ingalls.Standards %>%
  filter(Compound.Name %in% test.RFratios$Compound.Name) %>%
  rename(RF.ratio = QE.RF.ratio) %>%
  select(Compound.Name, RF.ratio) %>%
  mutate(Compound.Name = as.character(Compound.Name)) %>%
  mutate(RF.ratio = as.character(RF.ratio))

transect.RFratios <- transect.RFratios %>%
  as.data.frame() %>%
  filter(!is.na(RF.ratio)) %>%
  rbind(test.standards)
```


Switch to BMIS'd transect dataset.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
Samp.Data.transect <- read.csv("data_processed/BMIS_Output.csv") %>%
  separate(Run.Cmpd, sep = " ", into = c("Sample.Name")) %>%
  mutate(Compound.Name = MassFeature) %>%
  filter(Compound.Name %in% transect.RFratios$Compound.Name) %>%
  left_join(transect.RFratios) %>%
  left_join(transect.RFs.dimensions %>% select(Compound.Name, Column, RF.max, RF.min) %>% unique(), by = "Compound.Name") %>%
  mutate(RF.ratio = as.numeric(RF.ratio)) %>%
  select(Compound.Name, FinalBMIS, Sample.Name, Adjusted_Area, everything())  
```


*Volume of seawater filtered for all TRANSECT samples was 5 L*

Calculate umol in vial for eddy center compounds without an internal standard.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
Quan.Dat.transect <- Samp.Data.transect %>%
  mutate(RF.ave = as.numeric(rowMeans(Samp.Data.transect[, c("RF.min", "RF.max")]))) %>%
  mutate(umol.in.vial.ave = Adjusted_Area/RF.ave/RF.ratio,
         umol.in.vial.max = Adjusted_Area/RF.min/RF.ratio,
         umol.in.vial.min = Adjusted_Area/RF.max/RF.ratio) %>%
  select(Compound.Name:Adjusted_Area, everything())
```


Pull out data for eddy center Internal Standard and matched ones.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
original.IS.key <- read.csv("data_extras/InternalStandardNames.csv") %>%
  rename(FinalBMIS = Internal_Standards) %>%
  mutate(FinalBMIS = as.character(FinalBMIS))

IS.key <- Samp.Data.transect %>%
  select(FinalBMIS, MassFeature) %>%
  rename(Compound.Name = MassFeature) %>%
  unique() %>%
  left_join(original.IS.key %>% select(FinalBMIS, Concentration_nM)) %>%
  filter(str_detect(FinalBMIS, as.character(Compound.Name)))

rm(original.IS.key)
```

*This section is a little hairy and needs to be checked and cleaned of redundancies. Also check the Isethionic acid that has that has 186 umol/vial. Another result of that low standard area.*

Calculate umol in vial for eddycenter compounds that have a matched internal standard.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
IS.data.transect <- HILICS.transect %>%
  filter(as.character(Compound.Name) %in% IS.key$FinalBMIS) %>%
  mutate(IS_Area = Area.with.QC,
         FinalBMIS = Compound.Name) %>%
  select(IS_Area, FinalBMIS, ReplicateName) %>%
  left_join(IS.key %>% select(FinalBMIS, Concentration_nM))

IS.names <- data.frame(Compounds = c(IS.key[ ,"FinalBMIS"], as.character(IS.key[ ,"Compound.Name"])))

IS.smp.data.transect <- HILICS.raw.transect %>%
  left_join(IS.data.transect %>% select(Compound.Name, Concentration_nM)) %>%
  unique() %>%
  filter(Compound.Name %in% IS.names$Compounds) %>%
  filter(!str_detect(ReplicateName, "Std")) %>%
  mutate(Std.Type = ifelse(str_detect(Compound.Name, ","), "Internal_std", "Standard")) %>%
  mutate(testcol1 = ifelse(str_detect(Compound.Name, ","), sapply(strsplit(Compound.Name, ","), `[`, 1), Compound.Name)) %>%
  mutate(Names = ifelse(str_detect(testcol1, "-"), sapply(strsplit(testcol1, "-"), `[`, 2), testcol1)) %>%
  mutate(Pairs = ifelse(!str_detect(Compound.Name, ","), Compound.Name, paste(Names, "IS", sep = "_"))) %>%
  select(-c("Pairs", "testcol1", "Run.Type")) %>%
  mutate(Area.with.QC = as.numeric(as.character(Area.with.QC))) %>%
  arrange(ReplicateName) %>%
  group_by(Names) %>%
  group_split()


IS.mid_frame <- lapply(IS.smp.data.transect, function(x) group_by(x, ReplicateName))

IS.mid_frame2 <- lapply(IS.mid_frame,
                  function(x)
                  mutate(x,
                         umol.in.vial_IS = (Area.with.QC[Std.Type == "Standard"] / Area.with.QC[Std.Type == "Internal_std"]) * (Concentration_nM[Std.Type == "Internal_std"]/1000)))

IS.smp.data.transect <- do.call(rbind, IS.mid_frame2) %>%
  filter(!str_detect(Compound.Name, ",")) %>%
  rename(Sample.Name = ReplicateName) %>%
  select(Sample.Name:Area.with.QC, Concentration_nM, umol.in.vial_IS)

rm(list = c("IS.names", "HILICS.raw.eddycenter", "IS.mid_frame", "IS.mid_frame2"))
```


Add IS_smp info back into main frame.
*Add a test to make sure the sample names are what they should be*
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
all.info <- Quan.Dat.transect %>%
  left_join(IS.smp.data.transect %>% select(Sample.Name, Compound.Name, umol.in.vial_IS)) %>%
    mutate(umol.in.vial.ave = ifelse(is.na(umol.in.vial_IS), umol.in.vial.ave, umol.in.vial_IS),
         umol.in.vial.max = ifelse(is.na(umol.in.vial_IS), umol.in.vial.max, NA),
         umol.in.vial.min = ifelse(is.na(umol.in.vial_IS), umol.in.vial.min, NA)) %>%
  rename(ReplicateName = Sample.Name) %>%
  filter(!str_detect(ReplicateName, "DDA"))

```


Add in dilution factor from standard.info.

(nmol.in.Enviro.ave = umol.in.vial.ave * conversion factor * injection volume / Volume filtered * conversion factor * Dilution.Factor)
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
Dilution.Factor = 2
quanDat2 <- all.info %>%
  mutate(nmol.in.Enviro.ave = (umol.in.vial.ave*10^-6*400/5*1000 * Dilution.Factor)) %>% 
  left_join(HILICS.transect %>% select(Compound.Name, Emperical.Formula)) %>%
  select(Compound.Name, ReplicateName, Adjusted_Area, Orig_RSD:Emperical.Formula) %>%
  unique()

```

Obtain numbers for Carbon and Nitrogen according to their emperical formulas
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
# Okay dokay, go get how many Carbons and Nitrogens there are here.

quanDat3 <- quanDat2  %>%
  mutate(C = ifelse(is.na(str_extract(Emperical.Formula, "^C\\d\\d")),
                       str_extract(Emperical.Formula, "^C\\d"),
                      str_extract(Emperical.Formula, "^C\\d\\d"))) %>%
  mutate(C = as.numeric(str_replace_all(C, "C", ""))) %>%
  mutate(N = ifelse(str_detect(Emperical.Formula, "N\\D"),
                      1,
                      str_extract(Emperical.Formula, "N\\d"))) %>%
  mutate(N = as.numeric(str_replace_all(N, "N", ""))) %>%
  mutate(nmol.C.ave = nmol.in.Enviro.ave*C,
         nmol.N.ave = nmol.in.Enviro.ave*N ) %>%
  select(Compound.Name, SampID, ReplicateName, everything())

quanDatSum <- quanDat3 %>%
  group_by(Compound.Name) %>%
  summarise(nmol.Enviro.med = median(nmol.in.Enviro.ave, na.rm  = T),
            nmol.Enviro.min = min(nmol.in.Enviro.ave, na.rm  = T),
            nmol.Enviro.max = max(nmol.in.Enviro.ave, na.rm  = T),
            nmol.C.med = median(nmol.C.ave, na.rm  = T),
            nmol.C.min = min(nmol.C.ave, na.rm  = T),
            nmol.C.max = max(nmol.C.ave, na.rm  = T)) %>%
  arrange(desc(nmol.Enviro.med))

```

Organize final outputs and download.
```{r echo = FALSE, warning = FALSE, message = FALSE, error = FALSE}
# Cacluate mole fractions of each compound.
TotalMoles <- quanDat3  %>%
  select(SampID, nmol.C.ave, nmol.N.ave) %>%
  group_by(SampID) %>%
  summarise(totalCmeasured_nM_perID = sum(as.numeric(nmol.C.ave), na.rm = TRUE),
            totalNmeasured_nM_perID = sum(as.numeric(nmol.N.ave), na.rm = TRUE))

quanDat4 <- quanDat3 %>%
  unique() %>%
  left_join(TotalMoles) %>%
  mutate(ratioCN = totalCmeasured_nM_perID / totalNmeasured_nM_perID) %>%
  mutate(molFractionC = nmol.C.ave/totalCmeasured_nM_perID,
        molFractionN = nmol.N.ave/totalNmeasured_nM_perID) %>%
  select(Compound.Name, ReplicateName, Adjusted_Area, Area.with.QC, RF.ratio:molFractionN) %>%
  unique()

quanDatSum <- quanDat4 %>%
  group_by(Compound.Name) %>%
  summarise(nmol.Enviro.med = median(nmol.in.Enviro.ave, na.rm  = T),
            nmol.Enviro.min = min(nmol.in.Enviro.ave, na.rm  = T),
            nmol.Enviro.max = max(nmol.in.Enviro.ave, na.rm  = T),
            nmol.C.med = median(nmol.C.ave, na.rm  = T),
            nmol.C.min = min(nmol.C.ave, na.rm  = T),
            nmol.C.max = max(nmol.C.ave, na.rm  = T),
            mol.C.Fraction.med = median(molFractionC, na.rm = T),
            mol.C.Fraction.min = min(molFractionC, na.rm = T),
            mol.C.Fraction.max = max(molFractionC, na.rm = T)) %>%
  arrange(desc(Compound.Name))


write.csv(quanDatSum, "data_processed/Quantified_Summary.csv")
write.csv(quanDat4, "data_processed/Quantified_Full.csv")
write.csv(TotalMoles, "data_processed/Quantified_per_SampID.csv")

```



